---
title: "Introduction to PepSeq"
author: "Derek Sonderegger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`PepSeq` is a package developed to facilitate the exploration of PMI's PepSeq data.  Given a pulldown, we want to explore how the cleaved-uncleaved signal varies across different locations in the proteins. 

To begin, we will install the package via the following R commands:
```{r, eval=FALSE}
library(devtools)
install_github('dereksonderegger/PepSeq')  # install latest version of Derek's PepSeq tools
```
```{r, message=FALSE, warning=FALSE}
library(tidyverse)   # load the usual set of data processing tools
library(PepSeq)      # load Derek's library of routines
```

The first step is to set the working directory to wherever the pulldown .csv file is located.  This can be done via the point and click interface (Session -> Set Working Directory) or using the command `setwd`. Aternatively, you could just list the full pathname to the input and output files.  
```{r}
input_file  <- '~/GitHub/PepSeq/inst/extdata/PepSeq_vs_NN.csv'
output_file <- '~/GitHub/PepSeq/inst/extdata/PepSeq_vs_NN.pdf'
```



`PepSeq` assumes that a pulldown .csv file is arrange in a format with one protein/location combination
per line and then several sets of cleaved/uncleaved pairs and possibly columns of data that aren't cleaved/uncleaved but rather just some other signal (e.g. output from another model as to the likelihood of binding).

```{r}
read.csv(input_file) %>% colnames()
```

Notice that in the example pulldown file, there is just one experiment (wTEV) and two other added columns model of neural net model predictions and if the number of literature citations.  The orderings of all the columns doesn't matter, but you need to be able to identify the protein and position columns and each of the cleaved/uncleaved pairs need to have the same prefix followed by an '_Cleaved' or '_Uncleaved'.  If there is no 'Cleaved' or 'Uncleaved' suffix, then we assume each column is a measurement and will be data to be visualized.

In order to allow there to be many other interesting columns that should be ignored in the visualization, `PepSeq` requires you to indicate which columns are to be plotted by either having an initial suffix indicating it is data or to indicate the responses by column location (columns 3 to 7).

```{r, warning=FALSE, message=FALSE}
plot_pulldown( file = input_file,                         # input file
               output_file = output_file,                 # output file name
               protein_column = 'Protein_ID',             # which is the protein name
               position_column = 'Start_Loc',             # which is the starting location
               read_indicator = 3:6 )                     # what column range are data to be visualized

plot_pulldown( file = input_file,                        
               output_file = output_file,                 
               protein_column = 'Protein_ID',             
               position_column = 'Start_Loc',             
               read_indicator = c(3,4,5,6) )              # or specify individual columns
```


After either of the above are run, a pdf has been created and you can easily view it.

To add peaks to the graph, we need to add another option

```{r}
plot_pulldown( file = input_file,                         # input file
               output_file = output_file,                 # output file name
               protein_column = 'Protein_ID',             # which is the protein name
               position_column = 'Start_Loc',             # which is the starting location
               read_indicator = 3:6,                      # what column range are data to be visualized
               peaks = TRUE,                              # Show the peaks
               peak_method = 'PoT',                       # Peak creation method is Peaks of Threshold
               peak_param = c(.5, 90, 10) )               # thresholds are by row 
```



